### 2025-06-05 correlation.qmd
20250605 - converting a covariance to a correlation

The covariance is an unstandardized index of the association between two variables.
However, it can be helpful to use a standardized index to examine the [effect size](#sec-practicalSignificance)—i.e., the strength of the association.
The [Pearson correlation coefficient](#sec-correlationExamplesPearson), $r$, is an example of a standardized index of the covariance between two variables.

>>>>

The covariance is an unstandardized index of the association between two variables.
However, it can be helpful to use a standardized index to examine the [effect size](#sec-practicalSignificance)—i.e., the strength of the association.
The [Pearson correlation coefficient](#sec-correlationExamplesPearson), $r$, is an example of a standardized index of the covariance between two variables.
We can convert a covariance to a [correlation](#sec-correlationExamplesPearson) by standardizing it; i.e., by dividing by the multiplication of the standard deviation of $x$ and the standard deviation of $y$, as in @eq-covToCor:

$$
r_{xy} = \frac{\text{Cov}(x, y)}{\sigma_x \cdot \sigma_y}
$$ {#eq-covToCor}

For instance, for the covariance between height and fantasy points, we can convert it to a [correlation](#sec-correlationExamplesPearson) (we subset to just those rows that have data on both $x$ and $y$ to ensure we compute the standard deviations using the same values that are included in the estimate of the covariance):

```{r}
complete_cases <- player_stats_seasonal[complete.cases(player_stats_seasonal$height, player_stats_seasonal$fantasyPoints),]

cov_value <- cov(
  complete_cases$height,
  complete_cases$fantasyPoints,
  use = "pairwise.complete.obs")

cor_value <- cov_value / (sd(complete_cases$height, na.rm = TRUE) * sd(complete_cases$fantasyPoints, na.rm = TRUE))

cov_value

cor_value

cor(
  complete_cases$height,
  complete_cases$fantasyPoints,
  use = "pairwise.complete.obs")
```
